Spatial distribution of environmental DNA in a nearshore marine habitat

### 2017-01-10 Manuscript/v1.1_PeerJ/carkeek_grid.tex
fix references

While rigorous laboratory studies have investigated the effects of some environmental factors on eDNA persistence \citep{Klymus2015, Barnes2014, Sassoubre2016} and the transport of eDNA in specific contexts \citep{Deiner2014}, we suggest that field studies comparing the spatial distribution of communities of eDNA with expectations based on prior knowledge of organisms' distributions are also critical to developing a working understanding of eDNA in the real world.
Research to date has documented the non-random spatial distribution of % spatial variation of
meiofaunal \citep{Fonseca2014,Guardiola2016}, microbial \citep{Lallias2015}, and extracellular \citep{Guardiola2015} eDNA of marine and estuarine sediments, and of microscopic plankton in open ocean waters \citep{DeVargas2015}.
These studies conducted targeted sampling at intermediate (tens of kilometers) to global (thousands of kilometers) scales.
Here, we use a grid-based environmental sampling strategy to assess spatial variability of eDNA in a coastal marine environment at a fine scale (tens to thousands of meters), using molecular methods that focus on macrobial metazoans.

@@

\subsection*{Filtration}
One liter from each water sample was filtered in the lab on a clean polysulfone vacuum filter holder fitted with a 47 millimeter diameter cellulose acetate membrane with 0.45 micrometer pores. Filter membranes were moved into 900 microliters of Longmire buffer \citep{Longmire1997} using clean forceps and stored at room temperature \citep{Renshaw2014}. To test for the extent of contamination attributable to laboratory procedures, we filtered three replicate 1 liter samples of deionized water. These samples were treated identically to the environmental samples throughout the remaining protocols.

\subsection*{DNA Purification}
DNA was purified from the membrane following a phenol:chloroform:isoamyl alcohol protocol following Renshaw \citep{Renshaw2014}. Preserved membranes were incubated at 65 ºC for 30 minutes before adding 900 microliters of phenol:chloroform:isoamyl alcohol and shaking vigorously for 60 seconds. We conducted two consecutive chloroform washes by centrifuging at 14,000 rpm for 5 minutes, transferring the aqueous layer to 700 microliters chloroform, and shaking vigorously for 60 seconds. After a third centrifugation, 500 microliters of the aqueous layer was transferred to tubes containing 20 microliters 5 molar NaCl and 500 microliters 100\% isopropanol, and frozen at -20 ºC for approximately 15 hours. Finally, all liquid was removed by centrifuging at 14000 rpm for 10 minutes, pouring off or pipetting out any remaining liquid, and drying in a vacuum centrifuge at 45 ºC for 15 minutes. DNA was resuspended in 200 microliters of ultrapure water. Four replicates of genomic DNA extracted from tissue of a species absent from the sampled environment (\textit{Oreochromis niloticus}) served as positive control for the remaining protocols.

>>>>

While rigorous laboratory studies have investigated the effects of some environmental factors on eDNA persistence \citep{Klymus2015, Barnes2014, Sassoubre2016} and the transport of eDNA in specific contexts \citep{Deiner2014}, we suggest that field studies comparing the spatial distribution of communities of eDNA with expectations based on prior knowledge of organisms' distributions are also critical to developing a working understanding of eDNA in the real world.
Research to date has documented the non-random spatial distribution of % spatial variation of
meiofaunal \citep{Fonseca2014,Guardiola2016}, microbial \citep{Lallias2015}, and extracellular \citep{Guardiola2015} eDNA of marine and estuarine sediments, and of microscopic plankton in open ocean waters \citep{DeVargas2015}.
These studies conducted targeted sampling at intermediate (thousands of meters) to global (thousands of kilometers) scales.
Here, we use a grid-based environmental sampling strategy to assess spatial variability of eDNA in a coastal marine environment at a fine scale (tens to thousands of meters), using molecular methods that focus on macrobial metazoans.

@@

\subsection*{Filtration}
One liter from each water sample was filtered in the lab on a clean polysulfone vacuum filter holder fitted with a 47 millimeter diameter cellulose acetate membrane with 0.45 micrometer pores. Filter membranes were moved into 900 microliters of Longmire buffer \citep{Longmire1997} using clean forceps and stored at room temperature \citep{Renshaw2015}. To test for the extent of contamination attributable to laboratory procedures, we filtered three replicate 1 liter samples of deionized water. These samples were treated identically to the environmental samples throughout the remaining protocols.

\subsection*{DNA Purification}
DNA was purified from the membrane following a phenol:chloroform:isoamyl alcohol protocol following Renshaw \citep{Renshaw2015}. Preserved membranes were incubated at 65 ºC for 30 minutes before adding 900 microliters of phenol:chloroform:isoamyl alcohol and shaking vigorously for 60 seconds. We conducted two consecutive chloroform washes by centrifuging at 14,000 rpm for 5 minutes, transferring the aqueous layer to 700 microliters chloroform, and shaking vigorously for 60 seconds. After a third centrifugation, 500 microliters of the aqueous layer was transferred to tubes containing 20 microliters 5 molar NaCl and 500 microliters 100\% isopropanol, and frozen at -20 ºC for approximately 15 hours. Finally, all liquid was removed by centrifuging at 14000 rpm for 10 minutes, pouring off or pipetting out any remaining liquid, and drying in a vacuum centrifuge at 45 ºC for 15 minutes. DNA was resuspended in 200 microliters of ultrapure water. Four replicates of genomic DNA extracted from tissue of a species absent from the sampled environment (\textit{Oreochromis niloticus}) served as positive control for the remaining protocols.
